---
title: 'Lab: Class One Survey Analysis'
output:
  html_document
---

# 1. Import class one survey data from our Github site (https://github.com/kijohnson/ADA-2024) (The dataset is called ‘Class 1 Survey Fall 2024_di.csv’), calling the R dataframe that you create C1survey.

```{r}
# install.packages("pacman")
pacman::p_load(haven, readr, dplyr, openxlsx)

# import data
C1survey <-
 read_csv(
   "https://raw.githubusercontent.com/kijohnson/ADA-2024/main/Class%201%20(Getting%20started%20and%20RR%20I)/Lab/Class%201%20Survey%20Fall%202024_di.csv"
   )
```

# 2. Determine: a. the number of observations (i.e. the number of people who filled out the survey) and b. the number of variables in the dataframe.

```{r}
# View C1survey
View(C1survey)

#Calculate the number of rows and columns
dim(C1survey)
```

# 3. a. Rename the column variables to something shorter and that is descriptive of what the variable is about (for example like_dogs for the ‘Do you like dogs?’ question variable) and b. Display that they are renamed.

```{r}
#use names function to rename the 27 variables in order
names(C1survey)[1:27]<-c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby", "program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length", "top_three","public_health_interest", "fav_num", "bday", "bmonth", "country", "state", "city", "highest_educ_level")

# check column names
colnames(C1survey)
```

# 4. Write code to determine and display the number of factor, integer, numerical, and character variables there are in the C1survey dataset.

```{r Determine variables}

# get a basic summary of variable types using the sapply function
table(sapply(C1survey, typeof))

```

# 5. a. Using code, check the bday and bmonth variables for any unusual or missing values. If you find any, b. describe what you will do with the unusual values in a comment before or immediately following the code that does it, and c. after you have cleaned up any unusual values, find the median bday and month. *Note: please make sure that your bday and month variables are numeric before doing statistics

```{r}
## Converting the character variable CALORIES to numeric

#a.
# first check class again
class(C1survey$bday)
class(C1survey$bmonth)

# convert variables to numeric using the as.numeric function for stats
C1survey$bday_n <-as.numeric(C1survey$bday)
C1survey$bmonth_n <-as.numeric(C1survey$bmonth)

# check the class of the new numeric variables
class(C1survey$bday_n)
class(C1survey$bmonth_n)

## simple statistics

# get summary stats
summary(C1survey$bday_n) 
summary(C1survey$bmonth_n)

#b.
#delete NAs
C1survey_clean <- na.omit(C1survey)
summary(C1survey_clean)

#c.
# get median
median(C1survey_clean$bday_n, na.rm=TRUE)
median(C1survey_clean$bmonth_n, na.rm=TRUE)

```

# 6. a. Create a new variable called bseason that gives the season according to Northern Meteorological season in which respondents were born (winter = 12/1 to 2/29, spring = 3/1 to 5/31, summer = 6/1 to 8/31, fall = 9/1 to 11/30). b. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct c. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code.

```{r}
# a.
# using dplyr
C1survey_clean <- C1survey_clean %>%
  mutate(bseason=case_when(bmonth %in% c(12,1,2) ~"Winter",
                           bmonth %in% c(3,4,5) ~"Spring",
                           bmonth %in% c(6,7,8) ~"Summer",
                           bmonth %in% c(9,10,11) ~"Fall"))

# check variable
table(C1survey_clean$bmonth, C1survey_clean$bseason) 

C1survey_clean$bseason_f<-factor(C1survey_clean$bseason, levels=c("Winter", "Spring", "Summer", "Fall"))

# b. 
table(C1survey_clean$bmonth, C1survey_clean$bseason)
table(C1survey_clean$bmonth, C1survey_clean$bseason_f)

# c.
tab<-addmargins(table(C1survey_clean$bmonth, C1survey_clean$bseason, useNA = "always"), 1)
tab
tab<-addmargins(table(C1survey_clean$bmonth, C1survey_clean$bseason_f, useNA = "always"), 1)
tab
```

# 7. Pick your favorite variable or variables to analyze, come up with a question you want to answer with that variable or variables, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. #’s).

```{r}
# Rank according to the number in birth season
C1survey_clean %>%
  count(bseason) %>%
  arrange(desc(n))
```

